\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-18 paper/paper.tex
Updates to paper.

\documentclass[10pt, preprint]{sigplanconf}
\usepackage{times,epsfig,endnotes,graphicx}
\usepackage{datetime, url, hyperref}

@@

% Paper number and no. of pages as author
\authorinfo{Paper \textbf{\#XX}}{NN pages}

@@

\begin{abstract}
The introduction of non-volatile directly attached memory in datacenters - in
the form of NVDIMMs and new memory technologies allows efficient data structures
to be built on non-volatile heaps that provide good performance coupled with
durability. However high availability remains a problem for these primarily
single machine solutions. While transactional interfaces built over RDMA capable
networking hardware offer a viable alternative, such hardware is often
unavailable in datacenters. Cyclone offers a solution to this problem by
providing transparent replication middleware that can be used with Intel's Non
Volatile Memory Library (NVML) to construct durable data-structures that map
keys to values.  Cyclone is designed around an optimized userspace networking
stack to reduce latency. Further, it can scale its consensus protocol across
cores and network ports to provide good throughput under load. We show that
Cyclone can replicate as many as 7 million updates a second using just 8 cores
(for the consensus protocol) and 4 commodity 10 GigE ports on each replica.
\end{abstract}

@@

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients. The server side machines are each equipped with four commodity
82599-based 10 Gbit ethernet ports, while each client machine has one
82599-based 10 Gbit ethernet port. All the machines are connected to the same 10
Gbit switch. We use 8 cores on the server machines for running 8 instances of
RAFT and 39 cores to run application code. The last remaining core is reserved
for management code. We use NVML with the non volatile pool held in DRAM, which
proxies for the performance we would expect from NVDIMMs. We evaluate Cyclone in
three stages.

First, we evaluate pure replication by having Cyclone simply replicate the
contents of the RPC call without executing it and return a response to the
client when replication is complete.  Figure~\ref{fig:pure_rep} shows the
results. We consider two cases, one where the clients send an RPC call with no
payload and one where they add a 512 byte dummy payload. For the no payload case
the bottleneck is the message handling rate. We note that Cyclone can handle
small message replication rates as high as \emph{7 million messages a
  second}. In the case of RPC calls with 512 byte payloads, the bottleneck is
the available network bandwidth. To show this, we also plot aggregate transmit
throughput derived from these results by adding the rate at which data is sent
to replicas from all the 8 quorum leaders to the rate at which responses are
sent back to the clients.  With 3 replicas Cyclone shows \emph{slightly more
  than 40 Gbps} of aggregate transmit bandwidth. The reason for this is that the
quorum leaders can end up residing on different machines. Although replication
traffic is still limited by the receive bandwidth at follower replicas, the
responses to the clients, originating from different machines have access to
independent 4x10GigE ports, allowing total transmit throughput to exceed 40
Gbps. Although not discussed in this paper, Cyclone's control plane is designed
to exploit this by trying to spread quorum leaders across machines.

Also worth noting is that thanks to DPDK's low latency access to the NICs, three
way replication costs as little as 25 us of latency in the unloaded case. This
latency is measured at the client and is therefore four commodity 10 GigE
network hops through the switch, with each hop taking under 7 us including
network stack delays. These results should serve to illustrate that simple
commodity ethernet can indeed provide reasonable replication performance with a
carefully designed software stack that can exploit it.

\begin{figure}
\begin{tabular}{cc}
\begin{minipage}{0.25\textwidth}
  \hspace{-0.17in}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_mops_0.pdf}
\end{minipage}&
\begin{minipage}{0.25\textwidth}
  \hspace{-0.37in}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_gbps_0.pdf}
\end{minipage}\\
\begin{minipage}{0.25\textwidth}
  \hspace{-0.17in}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_mops_512.pdf}
\end{minipage}&
\begin{minipage}{0.25\textwidth}
  \hspace{-0.37in}
  \includegraphics[width=\textwidth,height=3cm]{results/replication_gbps_512.pdf}
\end{minipage}
\end{tabular}
\caption{Evaluation of pure replication performance}
\label{fig:pure_rep}
\end{figure}

\begin{figure}[htb]
%\vspace{-0.2in}
\centering
\begin{tabular}{c}
\includegraphics[height=3cm,width=0.8\columnwidth]{results/cc_mops.pdf}
\end{tabular}
\vspace{-0.22in}
\caption{Evaluation of replication performance with crash consistency}
\label{fig:cc_rep}
\vspace{-0.08in}
\end{figure}

\begin{figure}[htb]
%\vspace{-0.2in}
\centering
\begin{tabular}{c}
\includegraphics[height=3cm,width=0.8\columnwidth]{results/app_mops.pdf}
\end{tabular}
\vspace{-0.22in}
\caption{Evaluation of in-memory hash table performance with 3-way crash consistent replication}
\label{fig:app_rep}
\vspace{-0.08in}
\end{figure}

Next, Figure~\ref{fig:cc_rep} shows the effect of adding a crash consistent
transaction where the application still does nothing but Cyclone durably
remembers the last sequence number and RPC call result (the application simply
returns the received payload). Throughput is now bottlenecked by NVML allocator
performance at just under a million messages a second. The final result is
adding the hash table in the picture with the clients updating the values of
keys already present in the hash table - a stable workload for easy
comparison. As Figure~\ref{fig:app_rep} shows, performance stays at just under
1M calls a second for pure updates. We also show a pure read only case where the
clients only query the value of a key - there is no replication \emph{nor} any
crash consistent transaction applied. It is interesting to note that the pure
read only case is bottlenecked almost at the same point as the pure replication
case of 7 million operations a second, demonstrating the efficiency of Cyclone's
replication under load.

Contrary to common wisdom that replication over commodity networks is expensive,
these results lead us to conclude that Cyclone provides adequate performance
headroom for transparent replication in NVML-based key value maps.
%In fact our current focus in the light of
%these results is towards improving application performance on top of NVML rather
%than improving the performane of replication in Cyclone.
%NVML is a general purpose library written for a variety of
%substrates including traditional SSD based storage and there is adequate room
%for improvement of performance for applications running on directly attached
%persistent memory with close to DRAM performance.

>>>>

\documentclass[10pt, preprint]{sigplanconf}
\usepackage{times,epsfig,endnotes,graphicx}
\usepackage{datetime, url, hyperref, pdfpages}

@@

% Paper number and no. of pages as author
\authorinfo{Paper \textbf{\#224}}{NN pages}

@@

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions - with innovations targeted at taking
advantage of growing flash speeds. Unfortunately their performance is hampered
by the need to maintain and replicate a write ahead log to guarantee
availability in the face of machine and storage failures. Cyclone offers a
solution to this problem by leveraging a small amount of directly attached
non-volatile memory such as NVDIMMs to maintain a two level log - the upper
level in a small amount of NVDIMM memory draining into a larger lower level log
maintained on flash. Replicating the smaller log kept in NVDIMMs transforms the
replication problem into a packet switching problem - the leader sends the exact
same packet to all follower replicas and we solve the problem as such by using
the RAFT protocol implemented as a software multicast module on the Data Plane
Development Kit (DPDK). Second, we levergage the observation that key value store
interfaces are commutative for operations to different keys to scale our two
level log horizontally into a number of physical two level logs - with a hash of
the key being used to choose the right log. Each log is replicated by an
independent instance of RAFT, thereby leveraging the scalability afforded by
classic software packet switching when using multiple CPU cores. Cyclone is able
to replicate millions of small updates a second using only commodity 10 gigabit
ethernet adapters and improves the performance of RocksDB - a popular persistent
key value store used in industry by orders of magnitude when compared to its own
write ahead log without replication.
\end{abstract}

@@

\section{Evaluation}
We evaluate Cyclone on a 12 node cluster connected via a 10 GigE switch. Three
of the machines are equipped with 1.6TB Intel DC P3600 SSDs and 4*10 GigE
ports. The remaining nine machines do not have SSDs and have only one 10 GigE
port, serving as clients for most of the experiments. As with other
work~\cite{faast}, we use DRAM on the machines to proxy for NVDIMMs where
necessary - the persistent memory needed never exceeds 64 MB regardless of the
size of the key value store or second level log on flash. We divide the
evaluation into three parts. First, we evaluate Cyclone's performance with a
single level log as a pure software packet switch. Next, we evaluate performance
when adding a second level of log on flash. Finally, we evaluate performance
when integrated with Rocksdb as an alternative to Rocksdb's write ahead
log. Unless otherwise mentioned, we use a 60 byte header followed by a payload
for experiments. We log both the header and payload. In all cases the server
echoes the received entry (header and payload) back to the client. We fix the
number of application threads at 32 using at most 8 threads for running RAFT
instances.

Cyclone is built on top of DPDK to apply software packet switching techniques to
the log replication problem. We therefore begin by systematically evaluating
optimizations applied in Cyclone to replicate the top level log in
Figure~\ref{fig:network_opts} - with no payload. The y-axis reports latency seen
at the client (which means two network round trips with replication). Using
TCP/IP to replicate a RAFT log tops out at around 30K entries/s. Applying
optimizations improves performance by a large enough amount that we use a log
scale on the x-axis. Switching to DPDK (the line marked +DPDK) improves the
throughput by an order of magnitude to around 50K entries/s. Using batching
(line marked +batching) improves the performance further bringing us close to a
million entries/s. Scaling horizontally to 8 physical logs (+8 phy logs)
improves performance to close to 2M entries/s. Finally using all 4 ports on the
machine to replicate entries improves performance considerably to 6M
entries/s. In all performance improves by 200X over the TCP/IP single log
baseline. Cyclone also considerably improves the latency for replication, from
close to 100us with TCP/IP to around 30us at peak throughput.

\begin{figure}
\includegraphics[scale=0.6]{results2/network_opts.pdf}
\caption{Network optimizations for top level log}
\label{fig:network_opts}
\end{figure}

There are two factors that can have significant impact on Cyclone's
performance. First, the number of replicas dictates the outgoing message rate
from the leader replica and therefore increasing the replication factor can
decrease Cyclone's performance. Figure~\ref{fig:replicas} shows the impact of
varying replica count. Using only a single replica cuts out a network round trip
and shows the best unloaded latency (10 us) and peak throughput (near 10M
entries/s). Adding replicas decreases the peak throughput down to around 2M
entries/s. We note that a number of previous pieces of work~\cite{faast, farm}
use three replicas and therefore we focus on three replicas for the replicated
cases we consider below. The second factor that dictates Cyclone's performance
is the size of the log entry being replicated. Figure~\ref{fig:payload} shows
the effect of increasing the payload size from zero to 512 bytes. Peak
throughput drops from 6M entries/s to approximately 2M entries/s. At this
replication rate, the leader replica needs to transmit data at approximately 30
Gbit/s. Coupled with the cost of network headers all four 10 GigE ports are now
saturated and therefore Cyclone hits the network line rate bottleneck at this
point.

\begin{figure}
\includegraphics[scale=0.6]{results2/replicas.pdf}
\caption{Impact of replica count}
\label{fig:replicas}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/512.pdf}
\caption{Impact of payload size}
\label{fig:payload}
\end{figure}

We now turn our attention from the network component of Cyclone to the storage
one by adding the second level flashlog. We evaluate the impact of adding
storage to our hitherto pure packet switching scenario in
Figure~\ref{fig:flashlog}. Batching entries from the top level NVDIMM log to the
second level flashlog is clearly beneficial as adding flash storage at the
second level has almost no impact on peak performance in Cyclone for small
entries. The situation however changes for larger
entries. Figure~\ref{fig:flashlog_512} shows that using a 512 byte payload has a
significant impact on peak throughput - it drops to approximately 350K
ops/sec. This corresponds to around 50K 4KB IOPS to the SSD to write out the
flashlog pages. The peak for the drive is 160K IOPS using a queue depth
(concurrency) of 128. With 32 application threads we expect a lower peak
throughput around 40K IOPS explaining our bottleneck at 50K IOPS.  It is
possible to tune our observed performance further by aligning the flush boundary
to increase the number of outstanding requests - we do not do so in this paper,
keeping Cyclone agnostic to the exact size of entry being replicated. A final
point about Figure~\ref{fig:flashlog_512} is that once we are past the storage
bottleneck the latency spike is dramatic and large enough to trigger Cyclone's
failure detector and repeated retries from the clients. There are - therefore -
no points on the ``knee'' of the curve as in the pure packet switched one level
log case.

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog.pdf}
\caption{Impact of adding second level log}
\label{fig:flashlog}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/flashlog_512.pdf}
\caption{Impact of adding second level log (512 bytes payload)}
\label{fig:flashlog_512}
\end{figure}

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client reflecting the
unloaded case made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a rapid increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of quorums that must be synchronized
on. There is some amount of synchronization involved in the userspace DPDK
driver for access to common resources for accessing the NIC and this affects the
ability to simultaneously issues replication messages on all quorums. The
smaller contribution to increasing latency that continues past the 8 core case
is due to cacheline pingponging when executing the rendevous. Both these sources
of latency drift could be corrected using replication quorums mapped to
dedicated NICs and more scalable rendesvous designs (such as with machine aware
broadcast trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay, which would the minimum needed for a solution using
distributed transactions via two phase commit.

\begin{figure}
\includegraphics[scale=0.6]{results2/multi.pdf}
\caption{Ganged Operations}
\label{fig:ganged}
\end{figure}
